### 2017-11-30 reports/poster-agu-2017/blocks/conditional-simulation.tex
re-work directory

\begin{block}{Conditional Simulation}
  Goal: condition the RIP events (poorly represented in CM3) on EWD and MHC indices (better represented in CM3).
  Assume:
  \begin{itemize}
    \item Linear model for EWD and MHC indices
    \item MHC is secondary: if $\text{EWD}<0$, do not consider MHC
    \item Logistic regression model: $p(\text{RIP}_t) \sim f \qty(\text{EWD}_t, \text{MHC}_t)$
    \item Fit model on reanalysis temperature and pressure fields, and observed RIP events
    \item Simulate from model using CM3 temperature and pressure fields
  \end{itemize}
  \textbf{Comparing \cref{fig:historical-simulations} with \cref{fig:RIP-historic}, we see a marked improvement in model skill by conditioning on large-scale circulation features}
  \begin{figure}[ht]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/RIP_historical_sim_MAM.pdf}}
    \caption{
    (a) 10-year moving average of the number of MAM RIP days by year for the observed record (black solid line), the five conditional simulated RIP records.
    (b) Counts for the number of MAM RIP days by year for the observed record (black solid line), the five conditional simulated RIP records based on the GFDL CM3 ensemble member's geopotential height and temperature fields (light dashed lines), and the ensemble mean (thick dashed line).
    }
    \label{fig:historical-simulations}
  \end{figure}
\end{block}