### 2017-11-30 reports/poster-agu-2017/blocks/future-simulation.tex

\begin{block}{Future Simulations}
  \begin{itemize}
    \item A conditional simulation model using GCM atmospheric circulation and temperature projections for 21st century shows increasing trends for such events and
    \item Conditional simulation model attributes increasing RIP events approximately 1/3 to circulation changes and 2/3 to temperature changes
    \item Results differ signicantly from RIP events based only on MAM precipitation field
  \end{itemize}
  \begin{figure}[ht!]
    \centerline{\includegraphics[width=0.95\textwidth]{figs/RIP_future_sim_MAM.pdf}}
    \caption{
      (Left) 10-year moving average projections from 2006 to 2100 of MAM RIP counts by year from CM3 precipitation field (black solid line), conditional simulation based on CM3 temperature and pressure fields (light dashed lines show five simulations based on the one ensemble field and the dark dashed line shows the mean of these simulations).
      (Right top) Distribution of RIP counts by year for 2006-2050.
      (Right bottom) As (Right top) but for 2056-2100.
      }
      \label{fig:future-simulations}
  \end{figure}
\end{block}