### 2017-11-30 reports/poster-agu-2017/blocks/motivation.tex

\begin{block}{Conceptual Framework}
    \begin{itemize}
      \item Projections of future extreme precipitation remain highly uncertain \citep{IPCC2012}
      \item Rainfall (esp. extreme) poorly represented in GCMs \citep[\ie][]{Kendon2012}
      \item Precipitation fields often used for hydrological and decision models after bias correction \citep[e.g. quantile-quantile mapping][]{Gudmundsson2012}) is applied; however model deficiencies not identified, persistence characteristics in precipitation fields poorly represented \citep[e.g.][]{Piani2010}.
    \end{itemize}
    \emph{Is there a way to use credibly simulated state variables from GCM simulations to derive or simulate credible sequences of regional intense precipitation events associated with societal impacts?}
\end{block}