### 2017-11-30 reports/poster-agu-2017/blocks/summary.tex

\begin{block}{Summary}
  \begin{itemize}
    \item Recurrent atmospheric circulation patterns that correspond to regional intense precipitation events in  the Ohio River Basin are identified
    \item GCM biases in predicting regional intense precipitation frequency are largely overcome by conditioning these events on GCM atmospheric circulation patterns
    \item GCM frequency bias in RIP record appears a manifestation of inflated spatial correlation of high intensity precipitation
  \end{itemize}
  This \emph{framework for conditional simulation} is an alternative to grid-scale statistical downscaling
\end{block}