### 2017-11-30 reports/poster-agu-2017/blocks/thanks.tex

\begin{block}{Acknowledgements}
  This research was supported by funding from NSF Sustainability Research Networks (award 1444745): ``Integrated Urban Infrastructure Solutions for Environmentally Sustainable, Healthy, and Livable Cities'' and  SERDP grant \#2516: ``Climate Informed Estimation of Hydrologic Extremes for Robust Adaptation to Non-Stationary Climate''.
  The authors thank Yochanan Kushnir and James Booth for conversations regarding this work, and Casey Brown, Melissa Bukovsky, Rachel McCrary, Seth McGinnis, Linea Mearns, and Katherine Schlef for collaboration and advice.
\end{block}