### 2017-12-01 reports/poster-agu-2017/blocks/thanks.tex

\begin{block}{Acknowledgements}
  The authors thank David Farnham, Upmanu Lall, and Andrew Robertson for insightful conversations and guidance.
  JDG thanks the NSF GRFP program for support (grant DGE 16-44869).
  AGM was supported by the Atmospheric and Oceanic Sciences (AOS) Program at Princeton University.
\end{block}