### 2017-12-02 reports/poster-agu-2017/blocks/study-area.tex

\begin{block}{Lower Paraguay River Basin}
  Flat topography limits the river's ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time \cite{Bravo:2011et,Barros:2004bn}.
  During NDJF 2015-16, river stage [height] throughout the Lower Paraguay River Basin reached nearly three times climatological levels (not shown).
  \begin{mdframed}
  \begin{figure}
  	\noindent\includegraphics[width=0.9\textwidth]{study-area.jpg}
  	\caption{
  		Topographical map of the study area.
  	}
    \label{fig:study-area}
  \end{figure}
  \end{mdframed}
\end{block}