### 2018-03-01 reports/writeup/Supplement.tex
Update some more plots and work on writeup

%-------------------------------------------------------------------------------
% Preamble
%-------------------------------------------------------------------------------

\documentclass{article}

%Packages
\usepackage{
	siunitx, % units
  graphicx, % for images
	natbib,
	booktabs
}
\usepackage[letterpaper]{geometry}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\usepackage[breaklinks=true,hidelinks]{hyperref}
\graphicspath{{../../figs/}{./}}

%-------------------------------------------------------------------------------
% Title and Authors
%-------------------------------------------------------------------------------

\title{Supplemental Information for\\``Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and subseasonal-to-seasonal predictive skill''}
\author{James Doss-Gollin\and \'{A}ngel G. Mu\~{n}oz  \and Simon J. Mason \and Max Past\'{e}n }
\date{\today}

\begin{document}

%% Necessary!
\maketitle

This document contains several supplemental figures described in the main text.
Further supporting information is available in the form of source code and \texttt{jupyter} notebooks at \url{github.com/jdossgollin/PYFloods}.

\listoftables
\listoffigures

\clearpage

\begin{table}
	\centering
	\begin{tabular}{lrr}
		\toprule
		{} &  NDJF 2015-16 &  climatology \\
		wtype &               &              \\
		\midrule
		1     &      0.272727 &     0.212849 \\
		2     &      0.231405 &     0.184517 \\
		3     &      0.107438 &     0.156342 \\
		4     &      0.206612 &     0.156189 \\
		5     &      0.132231 &     0.149665 \\
		6     &      0.049587 &     0.140438 \\
		\bottomrule
		\end{tabular}
	\caption{Weather type (WT) occurrence fraction during NDJF 2015-16 and Climatology}
\end{table}

\clearpage

\begin{figure}
	\centering
  \includegraphics[width=\textwidth,height=0.6\textheight,keepaspectratio=true]{gpw-v4-2015.png}
	\caption{
		Gridded estimate of population density (color; in units of persons per square kilometer) from \citet{GPWv4}.
	}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{wt_classifiability.pdf}
	\caption{
		Classifiability index \citep[see Methods and][]{Michelangeli1995} calculated for several chosen values of $K$, representing the number of weather types produced.
		Results are produced with only 75 simulations for each $K$.
	}
\end{figure}

\begin{figure}
	\includegraphics[width=\textwidth]{nino_sst_anomalies.pdf}
	\caption{
		Monthly SST anomalies during major El Ni\~{n}o events.
		Months shown are (a) December 1982, (b) December 1997, and (c) December 2015.
		Units are in \si{\celsius}.
	}
\end{figure}

\clearpage
\bibliographystyle{ametsoc2014}
\bibliography{library}

\end{document}