### 2018-09-12 reports/poster/blocks/mjo-enso.tex

\begin{block}{ENSO, MJO, and Weather Types}
    \begin{framed}
        \begin{figure}
            \centering
            \caption{
                Anomalous probability of occurrence of each weather type concurrent with observance of each MJO and ENSO phase.
                Only values which are significant at $p < 0.10$ are shown.
            }\label{fig:wt-mjo-ensoflooded}
            \includegraphics[width=\textwidth]{wt_mjo_enso.pdf}
        \end{figure}
    \end{framed}
    \gls{wt} 1 occurs more frequently during El Ni\~{n}o years for most \gls{mjo} phases, particularly during phase 2.
    During El Ni\~no years, \gls{wt} 3 - associated with dryness over the \gls{lprb} - occurs less frequently during \gls{mjo} phases 4, 6, and 7, and more often during \gls{mjo} phase 8; this is consistent with the lack of \gls{wt} 3 during December 2015 and the frequent \gls{wt} 3 occurrence in mid-January 2016 (\cref{fig:fig:rain-wtype}).
\end{block}