### 2018-09-12 reports/poster/blocks/study-area.tex

\begin{block}{\acrlong{lprb}}
    Flat topography limits the river’s ability to carry the summer runoff, causing seasonal inundation of the Pantanal and distributing the river discharge in time~\cite{Barros:2004wg, Bravo:2012iq}.
    During NDJF 2015-16, river stage [height] throughout the Lower Paraguay River Basin reached nearly three times climatological levels (not shown).
    \begin{framed}
        \begin{figure}
            \noindent\includegraphics[width=0.7\textwidth]{study_area.jpg}
            \caption{
                Topographical map of the study area.
                Colors indicate $\log_{10}$ of elevation, in \si{\meter}.
                (a): all of South America.
                The domains of the \acrlong{lprb} and the domain used for weather typing are indicated in red and blue, respectively.
                (b): The \acrfull{lprb}.
                As for (a), the \gls{lprb} is marked with a red box.
            }\label{fig:study-area}
        \end{figure}
    \end{framed}
\end{block}