### 2018-09-12 reports/poster/poster.tex

\documentclass[serif,mathserif,final]{beamer}
\mode<presentation>{\usetheme{Lankton}}

\usepackage[
  orientation=landscape,
  width=40in,height=30in,%size=a0, % paper size
  scale=1.05 % font scale factor
  ]{beamerposter}
 \setbeamertemplate{caption}[numbered] % number figs

% Graphics Path
\usepackage{graphicx}
\graphicspath{{../../figs/}{figs/}}

% OTHER PACKAGE CALLS
\usepackage[utf8]{inputenc}
\usepackage{
  amsmath,amsfonts,amssymb,
  xspace,
  framed,
  siunitx,
  nth,
  physics
}
\sisetup{round-mode = figures, round-precision = 3}
\usepackage{cleveref}

%------ USE BIBLATEX WITH EXTREMELY MINAMALIST STYLE -----
% Use BibLaTeX
\usepackage[
	backend=biber,
	url=false,
	isbn=false,
	doi=false,
	maxnames = 2,
	style=numeric-comp,
	uniquename=false,
	uniquelist=false,
	firstinits=true,
	sorting=none,
	natbib=false,
	isbn=false,
]{biblatex}
\addbibresource{../library.bib}
\renewbibmacro{in:}{}

% One-paragraph bibliography environment
\defbibenvironment{bibliography}
{\list%
  {\printtext[labelnumberwidth]{%
    \printfield{prefixnumber}%
    \printfield{labelnumber}}%
  \ifentrytype{article}{}{}}%
    {\setlength{\leftmargin}{0pt}%
    \setlength{\topsep}{0pt}}%
    \renewcommand*{\makelabel}[1]{##1}}%
  {\endlist}%
{\mkbibitem}

% \mkbibitem just prints item label and non-breakable space
\makeatletter
\newcommand{\mkbibitem}{\@itemlabel\addnbspace}
\makeatother

% Add breakable space between bibliography items
\renewcommand*{\finentrypunct}{\addperiod\space}

% Set fields to clear
\AtEveryBibitem{%
	\clearfield{month}%
	\clearfield{day}%
  \clearlist {language}%why this has to be clear list.....
  \clearfield{pages}%
	\clearfield{pagetotal}%
	\clearfield{eprinttype}%
	\clearfield{eprint}%
	\clearfield{number}%
	\clearfield{volume}%
	\clearfield{issue}%
	\clearfield{name:given}
	\clearfield{name:first}
	\ifentrytype{article}{%
		\clearfield{title}%
	}%
}

%------- GLOSSARIES PACKAGE -------
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{wt}{WT}{weather type}
\newacronym{sallj}{SALLJ}{South American low-level jet}
\newacronym{enso}{ENSO}{El Ni\~{n}o-Southern Oscillation}
\newacronym{mjo}{MJO}{Madden-Julian Oscillation}
\newacronym{sacz}{SACZ}{South Atlantic Convergence Zone}
\newacronym{iri}{IRI}{International Research Institute for Climate and Society}
\newacronym{ecmwf}{ECMWF}{European Centre for Medium-Range Weather Forecasts}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}
\newacronym{eof}{EOF}{empirical orthogonal function}
\newacronym{mos}{MOS}{model output statistics}
\newacronym{xlr}{XLR}{homoscedastic extended logistic regression}
\newacronym{hxlr}{HXLR}{heteroscedastic extended logistic regression}
\newacronym{pcr}{PCR}{principal components regression}
\newacronym{cca}{CCA}{canonical correlation analysis}
\newacronym{sst}{SST}{sea-surface temperature}
\newacronym{scad}{SCAD}{South Central Atlantic Dipole}
\newacronym{lprb}{LPRB}{Lower Paraguay River Basin}

%-- Header and footer information ----------------------------------
\newcommand{\footleft}{Doss-Gollin, J., Á. G. Muñoz, S. J. Mason, and M. Pastén (2018), Heavy rainfall in Paraguay during the 2015-2016 austral summer: causes and sub-seasonal-to-seasonal predictive skill, Journal of Climate, doi:\href{dx.doi.org/10.1175/JCLI-D-17-0805.1}{10.1175/JCLI-D-17-0805.1}.}
\newcommand{\footright}{}
\title{Heavy Rainfall in Paraguay During the 2015-2016 Austral Summer: Causes and Sub-Seasonal-to-Seasonal Predictive Skill}
\author{James Doss-Gollin\inst{1} \quad \'{A}ngel G. Mu\~{n}oz\inst{2,3} \quad Simon J. Mason\inst{3} \quad Max Past\'{e}n\inst{4}}
\institute
{\inst{1} Columbia Water Center, Columbia University \inst{2} AOS, Princeton University \inst{3} International Research Institute for Climate and Society, Columbia University \inst{4} Direcci\'{o}n de Meteorolog\'{i}a e Hidrolog\'{i}a, Paraguay}

%-------------------------------------------------------------------

%-- Main Document --------------------------------------------------
\begin{document}
\begin{frame}{}
  \begin{columns}[t]

    %-- Column 1 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

        \input{blocks/at-a-glance.tex}
        \input{blocks/observations-wts.tex}

    \end{column}%1

    %-- Column 2 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

      \input{blocks/observed-anomalies.tex}
      \input{blocks/study-area.tex}

    \end{column}%2

    %-- Column 3 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

      \input{blocks/s2s-forecasts.tex}
      \input{blocks/model-output-stats.tex}

    \end{column}%3

    %-- Column 4 ---------------------------------------------------
    \begin{column}{0.22\linewidth}

      \input{blocks/mjo-enso.tex}
      \input{blocks/atlantic-pacific.tex}

      %-- Block 4-2
      \begin{block}{References}
          \renewcommand*{\bibfont}{\footnotesize}
          \printbibliography[heading=none]
      \end{block}

    \end{column}%1

  \end{columns}
\end{frame}
\end{document}